ANGSD-wrapper: utilities for analyzing next generation sequencing data

### 2016-01-29 manuscript.tex
Made additional edits to abstract
ill comment later
edit

\usepackage[colorinlistoftodos]{todonotes}
\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
\definecolor{amethyst}{rgb}{0.6, 0.4, 0.8}
%\setlength{\marginparwidth}{1.5cm}
\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
\newcommand{\ad}[1]{\todo[size=\scriptsize, color=blue]{#1}}
\newcommand{\ph}[1]{\todo[size=\scriptsize, color=pink]{#1}}
\newcommand{\tjyk}[1]{\todo[size=\scriptsize, color=green]{#1}}
\newcommand{\cl}[1]{\todo[size=\scriptsize, color=amethyst]{#1}}
\newcommand{\fst}{${F_{ST}}$ }          
\newcommand{\X}{\textcolor{red}{\bf X }}

@@

\begin{abstract}
High throughput sequencing has changed many aspects of population genetics, molecular ecology, and related fields, affecting both experimental design and data analysis.
The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
% \cl{added text}
ANGSD uses probabilistic approaches to calculate genome wide descriptive statistics.
The package is specifically designed to produce more accurate results for samples with low sequencing depth and makes use of full genome data while handling a wide array of sampling and experimental designs.
Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
% \cl{added text}
This software supports multiple types of analyses such as Site Frequency Spectrum, Genotype Likelihood estimations, Thetas diversity statistics and neutrality tests, Principal Component Analysis, Admixture, and Inbreeding coefficient calculations.
ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses. ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
We demonstrate the usefulness of ANGSD-wrapper by analyzing resequencing data from populations of wild and domesticated \textit{Zea}. 
ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
\end{abstract}

@@

We first use the ``SFS'' method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp. \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples. 
The SFS and diversity statistics were calculated using ancestral states inferred in the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively). Figure \ref{fig:figure3}C shows the distribution of pairwise nucleotide diversity in each window within our region of interest and each vertical line represents the number of windows with the corresponding value of $\pi$. 
Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.   
There are no genes in this region that have  been identified as potential domestication candidates, consistent with the lack of extended regions of high \fst in our analysis (Figure \ref{fig:suppfst}).

>>>>

\usepackage[colorinlistoftodos]{todonotes}
\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
\definecolor{amethyst}{rgb}{0.6, 0.4, 0.8}
%\setlength{\marginparwidth}{1.5cm}
\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
\newcommand{\ad}[1]{\todo[size=\scriptsize, color=blue]{#1}}
\newcommand{\ph}[1]{\todo[size=\scriptsize, color=pink]{#1}}
\newcommand{\tjyk}[1]{\todo[size=\scriptsize, color=green]{#1}}
\newcommand{\cl}[1]{\todo[size=\scriptsize, color=amethyst]{#1}}
\newcommand{\tvk}[1]{\todo[size=\scriptsize, color=razzledazzlerose]{#1}}
\newcommand{\fst}{${F_{ST}}$ }          
\newcommand{\X}{\textcolor{red}{\bf X }}

@@

\begin{abstract}
High throughput sequencing has changed many aspects of population genetics, molecular ecology, and related fields, affecting both experimental design and data analysis.
The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
% \cl{added text}
ANGSD uses probabilistic approaches to calculate genome wide descriptive statistics.
The package is specifically designed to produce more accurate results for samples with low sequencing depth and makes use of full genome data while handling a wide array of sampling and experimental designs.
Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
% \cl{added text}
This software supports multiple types of analyses such as Site Frequency Spectrum, Genotype Likelihood estimations, Thetas diversity statistics and neutrality tests, Principal Component Analysis, Admixture, and Inbreeding coefficient calculations.
%ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
We demonstrate the usefulness of ANGSD-wrapper by analyzing resequencing data from populations of wild and domesticated \textit{Zea}. 
ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
\end{abstract}

@@

We first use the ``SFS'' method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp. \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples. 
The SFS and diversity statistics were calculated using ancestral states inferred in the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively). Figure \ref{fig:figure3}C shows the distribution of pairwise nucleotide diversity in each window within our region of interest and each vertical line represents the number of windows with the corresponding value of $\pi$. \jri{no dude, not explaining what a histogram is, explain what the vertical line is (mean, median, etc.)}
Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.   
There are no genes in this region that have  been identified as potential domestication candidates, consistent with the lack of extended regions of high \fst in our analysis (Figure \ref{fig:suppfst}).
